---
title: "Untitled"
author: "Amanda Howarth"
date: "10/22/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rvest)
library(p8105.datasets)
```

## Strings & Factors 
strings- are exactly what they look like (characters on the screen)
factors- there is underlying structure that R uses in the background for factors 
- the words u see in R are lying on top of a numeric structure; and numeric structured is 1, 2, 3, 4, 5 ordered etc. 

strings - make sense for unstructured characterinfo or ALOT of information/ lots of words that are different from each other 
 
factors - make sense if u have variables that are descriptive with only a few categorical levels 
 - factors have a few meaningful levels that can be ordered 

read_csv will not treat anythign as a factor until u tell it to specifically 
read.csv will treat all characters as factors (not good)

## Strings 
you can do alot with strings. 
- can concatenate strings together (putting these two together)
 - could repace all M's with L's or something 
 
## FACTORS 
factors make a differnece when you are thinking about plot ordering and how u build ur tables 
- reference category is always the first one 
- be aware of the underlying ordering 
- you may need to relevel the order for nice plots, etc.
- forcats package is way to go for factoring needs (part of the tidyverse)

## String Manipulation 
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
## str_detect --> to detect the presence or absence of a pattern in a string 
## in my string vector tell me whether the string "jeff" exists 

str_detect(string_vec, "m")
# found in name and my (TRUE TRUE FALSE FALSE)
```

```{r}
## replace one observation with another 
str_replace(string_vec, "jeff", "Jeff")
str_replace(string_vec, "m", "M")
```

```{r}
#creating new string vector 
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")

str_detect(string_vec, "^i think")
#starts with "I think"

str_detect(string_vec, "i think$")
# end of line (dollar sign)
```
```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Bb]ush")
# start with capital B or lowercase B, but not all caps
# can also do a range from a-z 
```
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")
# you want to detect a number followed by any letter 
``` 

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
# want to find everything starting with 7 followed by any character and then followed by 11 (thus AA711 would not work bc no character between 7 and 11)
# the period matches any character at all 
```

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, pattern = "\\[")
# what u use if u actually want to find a bracket. or want to find a period 
# backslash twice to say u are looking for bracket character 
```

## use strings in practice

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_nodes(css = "table") %>% 
  .[[1]] %>%
  html_table() %>%
  slice(-1) %>%
  as_tibble()

data_marj = 
  table_marj %>%
  select(-contains("P Value")) %>%
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") %>%
  separate(age_year, into = c("age", "year"), sep = "\\(") %>%
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) %>%
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

# you could see percent was still being treated as a character vairable so treat it as numeric now 
```

```{r}

#below is in regards to a problem on the webpage that uses html but it took my R too long to download  
## you can change page number to 2 to read in the second page, right now it will just be the first page it will read in 

```

## FACTORS 
factors are the way to store categorical variables in R 
- you see the layer on top (like male and female) but underneath the variables R has laid a numeric structure 

```{r}

vec_sex = factor(c("male", "male", "female", "female"))
vec_sex

as.numeric(vec_sex)

vec_sex = relevel(vec_sex, ref = "male")
vec_sex

as.numeric(vec_sex)

```

```{R}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

